---
title: "Exploratory Data Analysis"
output: html notebook
---

We'll use a dataset called "diamonds"
```{r}
library(ggplot2)
library(tibble)
library(dplyr)
```

```{r}
?diamonds
diamonds
```

```{r}
glimpse(diamonds)
```

```{r}
summary(diamonds)
```

```{r}
dim(diamonds)
```

```{r}
library(skimr)
skim(diamonds)
```

```{r}
View(diamonds)
```

## Visualize distributions
### Categorical Variables
```{r}
ggplot(data = diamonds)+
  geom_bar(mapping = aes(x=cut))
```

```{r}
diamonds %>% count (cut)
```

```{r}
table(diamonds$cut)
#diamonds $cut
```

```{r}
unique(diamonds$cut)
```

```{r}
levels(diamonds$clarity)
```

Compare two categorical values
```{r}
table(diamonds$cut, diamonds$clarity)
```

```{r}
diamonds %>% 
  count(cut, clarity) %>%
  arrange(desc(n))
```

Plot two categorical values
```{r}
ggplot(diamonds, aes(x=cut, fill = clarity))+
  geom_bar(position = "dodge")+
  theme(axis.text.x = element_text(angle = 90))
```

```{r}
ggplot(diamonds, aes(x=cut, fill = clarity))+
  geom_bar(position = "dodge")+
  coord_flip()
```

```{r}
ggplot(diamonds, aes(x=cut, fill = clarity))+
  geom_bar(position = "fill")+
  ylab("Proportion")
```
Percentages
```{r}
prop.table(table(diamonds$cut, diamonds$clarity))
```
Bar chart for each clarity
```{r}
ggplot(diamonds, aes(x=cut))+
  geom_bar()+
  facet_wrap(~clarity)+
  theme(axis.text.x = element_text(angle = 90))
```

```{r}
round(prop.table(table(diamonds$cut, diamonds$clarity))*100,2)
```

Continuous variables
```{r}
ggplot(data = diamonds, mapping = aes(x=carat))+
  geom_histogram(binwidth = 0.5)
```

```{r}
diamonds %>%
  count(cut_width(carat, 0.5))
```

```{r}
smaller <- diamonds %>% 
  filter (carat<3)

ggplot(data = smaller, mapping = aes(x=carat))+
  geom_histogram(binwidth = 0.1)
```

Overlay multiple histograms
```{r}
ggplot(data = smaller, mapping = aes(x=carat, color = cut))+
  geom_freqpoly(binwidth = 0.1)
```

```{r}
ggplot(data = smaller, mapping = aes(x=carat))+
  geom_histogram(binwidth=0.01)
```

```{r}
ggplot(data = diamonds, mapping = aes(x=y), binwidth = 0.5)+
  geom_histogram()
```

```{r}
ggplot(data = diamonds, mapping = aes(x=y), binwidth = 0.5)+
  geom_histogram()+
  coord_cartesian(ylim = c(0,50))
```
Unusual diamonds
```{r}
diamonds %>% 
  filter(y<3 | y >20) %>%
  select(price, x, y, z) %>%
  arrange(y)
```
Missing values
```{r}
diamonds2 <- diamonds %>%
  filter(between(y, 3, 20))

ggplot(diamonds2)+
  geom_histogram(aes(x=y), binwidth = 0.5)
```

```{r}
diamonds %>%
  mutate(y = ifelse(y<3 | y >20, NA, y)) %>%
  ggplot(aes(x=x, y=y))+
  geom_point()
```
Same plot without removing the NA values
```{r}
diamonds %>%
  ggplot(aes(x=x, y=y))+
  geom_point()
```
Covariation
Categorical vs Continuous Variables
Use frequency plots
```{r}
ggplot(diamonds, aes(x=price))+
  geom_freqpoly(aes(color=cut), binwidth=500)
```
Density plot
```{r}
ggplot(diamonds, aes(x=price, y = ..density..))+
  geom_freqpoly(aes(color=cut), binwidth=500)
```
Boxplot
```{r}
ggplot(diamonds, aes(cut, price))+
  geom_boxplot()+
  coord_flip()
```

```{r}
ggplot(diamonds)+
  geom_count(aes(cut, y=color))
```

```{r}
diamonds %>%
  count(color, cut) %>%
  ggplot(aes(x=color, y=cut))+
  geom_tile(aes(fill=n))
```
Continuous vs Continuous
```{r}
ggplot(diamonds)+
  geom_point(aes(carat, price), alpha=1/50)
```

```{r}
ggplot(smaller, aes(carat, price))+
  geom_bin2d()
```

