## Simple linear regression

Source: https://www.geeksforgeeks.org/simple-linear-regression-using-r/

### Assumptions of Linear Regression:
1. Linear Relationship between the features and the target
2. Small or no multicollinearity between the features
3. Homoscedcity
4. Normal distribution of error terms
5. No autocorrelations 

```{r}
library(ggplot2)
```

```{r}
#dataframe
data <- data.frame(
    Years_exp = c(1.1,1.3,1.5,2.0,2.2,2.9,3.0,3.2,3.2,3.7),
    Salary = c(39343,46205,37731,43525,39891,56642,60150,54445,64445,57189)
)
```

```{r}
View(data)
```

```{r}
colnames(data)
```

```{r}
plot(data$Years_exp, data$Salary,
    xlab = "Years of Experience",
    ylab = "Salary",
    main = "Scatter plot of salary against years of experience")
```

Find the regression line

Least Square technique

lm(Y ~ model)

```{r}
library(caTools)
split = sample.split(data$Salary, SplitRatio = 0.7)
trainingset = subset(data, split==TRUE)
testset = subset(data, split==FALSE)

# fitting simple linear regression to the training set
lm.r = lm(formula = Salary ~ Years_exp,
        data=trainingset)

# summary of the model
summary(lm.r)
```

### Predict values using the predict function
```{r}
# create a data frame with new input values
new_data <- data.frame(Years_exp = c(4.0,4.5,5.0))

# predict using the linear regression model
predicted_salaries <- predict(lm.r, newdata=new_data)

predicted_salaries
```

## Visualising the training dataset
```{r}
ggplot()+geom_point(aes(x=trainingset$Years_exp,
                    y = trainingset$Salary), color = "red")+
        geom_line(aes(trainingset$Years_exp,
        y= predict(lm.r, newdata=trainingset)), color= "blue")+
        ggtitle("Salary vs Experience (Training Set)")+
        xlab("Years of Experience")+
        ylab("Salary")
```

## Visualising the testing dataset
```{r}
ggplot()+
  geom_point(aes(x=testset$Years_exp, y = testset$Salary), color='red')+
  geom_line(aes(x=trainingset$Years_exp,
                y =predict(lm.r, newdata= trainingset)),
                color = 'blue')+
ggtitle("Salary vs Experience (Test set)")+
xlab("Years of Experience")+
ylab("Salary")
```