---
title: "Vectors Part 2"
output: html_notebook
---


```{r}
(x <- 1:10)
```
```{r}
seq(0, 10000, by=1000)
```

```{r}
y <- (11:20)
y
```

```{r}
x * y
```
------
Missing Values
------
```{r}
x <- c(0.5, NA, 3, 4)
```

Test if a vector has a missing value. Use "is.na"

Lists
```{r}
list(1, 3L, "a", TRUE)
```

```{r}
xlist <- list(data(iris))
xlist$data
```
df
```{r}
iris
```

```{r}
data <- iris3
data
```

Access all R datasets using data()

Creating dataframes

```{r}
mtcars
```

Create dataframe
```{r}
data<-data.frame(Name=c("John", "Bob"),
           Age=c(35,40), 
           Hair = c("brown", "black"),
           Married=c(TRUE, FALSE))
```

Helper functions
```{r}
df <- data.frame(id = letters[1:10], x=1:10, y=11:20)
```


```{r}
df
```

```{r}
head(df)
```

```{r}
str(mtcars)
```
```{r}
rownames(mtcars)
```


```{r}
colnames(mtcars)
```

Tibbles

```{r}
library(tibble)
```

```{r}
tb <- as_tibble(iris)
```

Intermediate R: relational operators
```{r}

```

Factors

Dates and Times
{lubridate}



```{r}
library(lubridate)
```

```{r}
today()
now()
```

```{r}
date1 <- mdy("March 25th, 1998")
date1
```

Creating date objects from numbers

```{r}
date2 <- dmy(25031998)
date2
```
Creating date time
```{r}
date_time <- ymd_hms("2024-12-25 20:38:41 EAT")
```

Functional Programming
lapply()
```{r}
rep(10,20)
```
```{r}
values <- lapply(rep(10,20), runif)
```


```{r}
values
```
```{r}
means <- lapply(values, mean)
means
```
unlist()-- flatten out into a vector

```{r}
unlist(means)
```

```{r}
car_names <- rownames(mtcars)
car_names
```
Measure the length of each of the names

```{r}
unlist(lapply(car_names, nchar))
```
Data Import and Export

{readr}
read_csv()
write_csv()

```{r}
library(readr)
```

```{r}
data <- read_csv("C:/Users/earls/Octavia Carbon/Octavia Carbon - General/3. Workspaces/1. Common/8. R&D/Mechanical/0. Kesses/Experimental Data/Yields/Scripts/Vacuum Drying 06092024/data/6thSeptember2024.csv")
data
```
```{r}
write.csv(mtcars, "mtcars.csv")
```

web scraping

```{r}
library(httr)
```

```{r}
online_data <- GET("https://docs.google.com/spreadsheets/d/1QnJZ_2KHFev7EnzAV-G7X09c1jAK9IWd-GzqhIMd3fA/edit?gid=1403256836#gid=1403256836")
```
```{r}
content(online_data)
```
Working with Databases

```{r}

```

Data Manipulation in R

